---
title: "My First Post"
author: "Sheila Saia"
date: 2018-07-23T21:13:14-05:00
categories: ["R"]
tags: []
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

Welcome to the wateR blog! I've been considering starting a water-centric data science blog for a while now. Getting this blogdown website up and running has helped solidify my motivation to write regularly about water-related data science advances that are available to R users...and maybe Python users too (both languages are so useful!).

Some of the specific goals of this blog are to (in no special order):

*  Practice developing my science communication and R skills.<br>
*  Review and demo new packages, data sets, and methods that might be useful to water resources research.<br>
*  Highlight newly published studies that use data science to advance or could be applied to water resources research.<br>
*  Discuss the present and future of water data science.<br>
*  Interview experts about water-related topics.<br>


```{r cars}

```



```{r pie, fig.cap='A fancy pie chart.', tidy=FALSE}

```
